---
title: "CKME 136 Capstone - Retirement Income Adequacy"
author: "Melissa Cortina"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data and select attributes

```{r}
labourincomedata <- read.csv(file="C:/Users/cortinm/Documents/original_data.csv", header = TRUE)
```

There are a total of 147 attributs in this dataset. After reading through the associated literature for the dataset, 26 relevant were chosen. 

Select only the 26 columns:

```{r}
attributes <- c(7,8,10,11,14,17,23,26,29,50,51,55,57,58,61,69,72,73,74,76,79,89,114,120,122,140)
data <- labourincomedata[attributes]
#write.csv(data, file="C:/Users/huhmyman/Documents/selected_data")
```

## Observe dataset

```{r}
head(data)
tail(data)
summary(data)
str(data)
```

Check if there are any missing entries:

```{r}
apply(data, 2, function(x) sum(is.na(x)))
```

Since all results sum to 0, then there are no null entries in the dataset.


## Analysis of unanswered surveys

If an individual left an answer blank or refused to answer while filling out the survey, then the answers were "dummied" and were entered as "98", "99", etc. Using the dataset's literature, each attribute was analyzed to determine if the dummied entries would impact the results. If there are too many dummied entries, then the attribute could be removed. If the dummied entries are low, then the dummied entries were replaced with the mode of that particular attribute if it is a qualitative attribute. If it was a quantitative attribute, the dummied entries were replaced with the mean.

1) ecage26 - Person's age as of December 31  
No dummied fields

2) ecsex99 - Sex of respondent  
No dummied fields

3) marst26 - Marital status of person as of December 31  
71 dummied fields with "97"  
Mode = 1

```{r}
data$marst26[data$marst26==97]=1
summary(data$marst26)
```

4) mjacg26 - Person's major activity at the end of reference year (working, retired, etc.)  
No dummied fields

5) immst15 - Flag-Person is an immigrant  
34,905 dummied fields with "7"  
Attribute to be removed

6) pvreg25 - Province of residence  
387 dummied fields with "97"  
Mode = 35

```{r}
data$pvreg25[data$pvreg25==97]=35
summary(data$pvreg25)
```

7) dwtenr25 - Ownership of dwelling  
7 dummied fields with "7"  
Mode = 1

```{r}
data$dwtenr25[data$dwtenr25==7]=1
summary(data$dwtenr25)
```

8) mortg25 - Flag-A mortgage on the dwelling  
9,654 dummied fields with "9"  
Mode = 1  

```{r}
data$mortg25[data$mortg25==9]=1
summary(data$mortg25)
```

9) multj28 - Flag-Multiple job holder, any mth refyr  
6,861 dummied fields with "9"  
Mode = 2

```{r}
data$multj28[data$multj28==9]=2
summary(data$multj28)
```

10) alhrp28 - Total hours paid all jobs during refyr  
290 dummied fields with "9997"  
6,861 dummied fields with "9999"

```{r}
data$alhrp28[data$alhrp28==9997]=9999
mean.alhrp28 <- mean(data$alhrp28[!data$alhrp28==9999])
data$alhrp28[data$alhrp28==9999]=mean.alhrp28
summary(data$alhrp28)
```

11) mtlswk28 - Number of months since person last worked  
2,213 dummied fields with "997"  
35,700 dummied fields wtih "999"  
Attribute to be removed

12) yrxfte11 - Number of years of experience, full-yr full-time  
25 dummied fields with "98"  
4,685 dummied fields with "97"

```{r}
data$yrxfte11[data$yrxfte11==98]=97
mean.yrxfte11 <- mean(data$yrxfte11[!data$yrxfte11==97])
data$yrxfte11[data$yrxfte11==97]=mean.yrxfte11
summary(data$yrxfte11)
```

13) clwkr1 - Class of worker for this job  
15,719 dummied fields with "99"  
Mode = 1

```{r}
data$clwkr1[data$clwkr1==99]=1
summary(data$clwkr1)
```

14) prmjb1 - Flag to indicate if job is permanent  
4,579 dummied fields with "7"  
Change all to "9" = Not Applicable

```{r}
data$prmjb1[data$prmjb1==7]=9
summary(data$prmjb1)
```

15) fllprt1 - Flag-If job was full-time  
854 dummied fields with "7"  
20 dummied fields with "8"  
Change all to "9" = Not Applicable

```{r}
data$fllprt1[data$fllprt1==7]=9
data$fllprt1[data$fllprt1==8]=9
summary(data$fllprt1)
```

16) nocg2e6 - Occupational code  
1,465 dummied fields with "97"  
Change all to "99" = Not Applicable

```{r}
data$nocg2e6[data$nocg2e6==97]=99
summary(data$nocg2e6)
```

17) imphwe1 - Implicit hourly wage for this paid worker job  
19,893 dummied fields with "999.99"  
Kept unchanged

18) penpln1 - If person is covered by a pension plan connected with this job  
688 dummied fields with "7"  
11 dummied fields with "8"  
19,855 dummied fields with "9"  
Mode = 2

```{r}
data$penpln1[data$penpln1==7]=2
data$penpln1[data$penpln1==8]=2
data$penpln1[data$penpln1==9]=2
summary(data$penpln1)
```

19) uncoll1 - If person was a member of a union/collective agreement  
434 dummied fields with "7"  
5 dummied fields with "8"  
19,855 dummied fields with "9"  
Mode = 3

```{r}
data$uncoll1[data$uncoll1==7]=3
data$uncoll1[data$uncoll1==8]=3
data$uncoll1[data$uncoll1==9]=3
summary(data$uncoll1)
```

20) n07c3g10 - Industry code of employer  
331 dummied fields with "97"  
Change to "99" = Not Applicable

```{r}
data$n07c3g10[data$n07c3g10==97]=99
summary(data$n07c3g10)
```

21) pubpv10 - If employer is in the public or private sector  
No dummied fields

22) cqpc42 - CPP/QPP contributions  
No dummied fields

23) rppc42 - Registered pension plan (RPP) contributions / employer-sponsored  
No dummied fields

24) udpd42 - Union dues (oth professional premiums)  
No dummied fields

25) wgsal42 - Wages and salaries, before deductions (employment income)  
No dummied fields

26) hleveg18 - Highest level of education  
106 dummied fields with "97"
Change to "1" = Never attended school

```{r}
data$hleveg18[data$hleveg18==97]=1
summary(data$hleveg18)
```

## Boxplot analysis

The quantitative attributes are in the follwowing positions: 1, 10, 11, 12, 17, 22, 23, 24, 25

Boxplots for each attribute:

```{r}
quant <- c(1,10,11,12,17,22,23,24,25)
boxplot(data[1])
boxplot(data[10])
boxplot(data[11])
boxplot(data[12])
boxplot(data[17])
boxplot(data[22])
boxplot(data[23])
boxplot(data[24])
boxplot(data[25])
```

## Correlation analysis - quantitative

```{r}
dataquant <- data[quant]
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
{
    usr <- par("usr"); on.exit(par(usr)) 
    par(usr = c(0, 1, 0, 1)) 
    r <- abs(cor(x, y)) 
    txt <- format(c(r, 0.123456789), digits=digits)[1] 
    txt <- paste(prefix, txt, sep="") 
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 
    test <- cor.test(x,y) 
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " ")) 
    text(0.5, 0.5, txt, cex = cex * r) 
    text(.8, .8, Signif, cex=cex, col=2) 
}
pairs(dataquant, lower.panel=panel.smooth, upper.panel=panel.cor)
```


## Removal of attributes

Results from the above "dummy" analysis conclude that the following attributes should be removed from the dataset:  
immst15  
mtlswk28

From the correlation analysis for the quantitative attributes, the correlation between any two attributes is not strong. Therefore, no further removal of attributes is needed.

The elimination of these attributes will be performed using WEKA.


## Export dataset

```{r}
write.csv(data, file="C:/Users/cortinm/Documents/selecteddata.csv")
```
